---
title: "MachineLearningProject"
author: "Alan Yun"
date: "Sunday, April 26, 2015"
output: html_document
---

##Executive Summary
Purpose of this project is to develop the model to predict the classes for lifting barbell correctly and incorrectly in 5 different ways, using data from accelerometers on the belt, forearm, arm, and dumbell of 6 participatns. 
 

##Data Loading 
First of all, training data and testing data have been loaded

```{r dataload, echo=TRUE, results="hide"}
library(knitr)
library(randomForest)
library(caret)
library(grDevices)

#GetWorkingDirectory
wd <- getwd() 
if (!file.exists("./data")) {dir.create("./data")}
wd <- paste(wd,"/data", sep="")

#download file
fileUrl.training <-  "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
fileName.training <- "pml-training.csv"

fileUrl.testing <-  "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
fileName.testing <- "pml-testing.csv"

download.file(fileUrl.training, file.path(wd, fileName.training ), mode="wb", method="auto")
download.file(fileUrl.testing, file.path(wd, fileName.testing ), mode="wb", method="auto")

#read data
training <- read.csv(file.path(wd, fileName.training))
testing <- read.csv(file.path(wd, fileName.testing))
```

As a part of exploratory analysis, reviewed a number of dimension and variables
```{r data explore, echo=TRUE}
#explore data - observations and variables
dim(training)
```
From training data set, 19,622 observations with 160 variables found.  

##Pre-Process
Performed pre-proces for training data in three different ways. 
1. Remove missing data from training data set
2. Remove near Zero-variance variables
3. Remove columns which are not used for the model
4. Perform Principal Components Analysis (PCA)
```{r preprocess, echo=TRUE, results="hide"}
#handle missing data 
training.NAs <- apply(training,2,function(x) {sum(is.na(x))})
training.NoNAs <- training[,which(training.NAs == 0)]

testing.NAs <- apply(testing,2,function(x) {sum(is.na(x))})
testing.NoNAs <- testing[,which(testing.NAs == 0)]

#handle near Zero-variance variables
nzvTraining <- nearZeroVar(training.NoNAs)
training.nzv <- training.NoNAs[-nzvTraining]

nzvTesting <- nearZeroVar(testing.NoNAs)
testing.nzv <- testing.NoNAs[-nzvTesting]

#Remove unnecessary and non-numeric columns
removeCol <- c("X","user_name","raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp","new_window")
training.ColRemoved <- training.nzv[,!(names(training.nzv) %in% removeCol)]
testing.ColRemoved <- testing.nzv[,!(names(testing.nzv) %in% removeCol)]

#PCA
preProc <- preProcess(training.ColRemoved[, -54], method = "pca", thresh = 0.99)
trainPC <- predict(preProc, training.ColRemoved[, -54] )
testPC <- predict(preProc, testing.ColRemoved[, -54] )
```

##Develop model
Create partition for data validation - 70% for training and 30% for validation. Using partitioned training data, create prediction model using random forests. 
```{r modelFit, echo=TRUE}
#Create Partition for predict model
inTrain = createDataPartition(y=training.ColRemoved$classe, p = 0.7, list = FALSE)
Train.Train<-training.ColRemoved[inTrain,]
Train.Testing<-training.ColRemoved[-inTrain,]

#set randome seed for reproducible
set.seed(1234)

RFModel <- randomForest(classe~., data=Train.Train, importance=TRUE)
RFModel
```

##Confusion Matrix
Using RF model developed using training data set, classe for validation data set is predicted. Predicted value is compated with its original classe and confusion matrix is produced as below. 
```{r cm, echo=TRUE}
# Run model for validation data
PredTesting <- predict(RFModel, newdata=Train.Testing)
#Confusion matrix
confusionMatrix(PredTesting, Train.Testing$classe)

```
According to the Confusion matrix, a model correctly predicts classe for 5865 observatrions and makes only 19 errors. An accuracy of model is 99.66%. 


